Investigating the applications and limits of single particle light scattering using limited detection schemes

### 2022-10-27 Optimizationnotes.tex
Add files via upload

This has already been achieved with single microsphere using dynamic light scattering measurements of the trapping beam; where Bar-Ziv et al utilised dynamic light scattering to characterise the trap's stiffness and anisotropy based on the Brownian motion of a trapped silica microsphere \cite{1}. They note that their method assumes that the silica scatters isotropically which allows them to place a single detection fibre at $75^\circ$ from the propagation axis. In the case where the trapped object scatters anisotropically - such as for an asymmetrical dimer - it is necessary to utilise multiple detection fibres to build a better picture of the trapped object's orientation based on the produced light scattering. However, for inverted microscope optical traps there is a significant engineering challenge of placing enough detection fibres at the correct angles within a close enough distance to the trapping focus to collect a clear signal. This means that any fibre based detection will have to rely on maximising the information provided by the light scattering signals collected at a limited number of detection fibres. Excluding the engineering challenge involved with ensuring proper coupling between the probe laser and the detection fibres, the primary concern is how much useful information can be

@@

\section*{Results}
The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap. The initial orientation was assumed as strictly vertical (in line with the beam propagation direction). The dimer's position orientation was recorded every $10 \mu s$ for using as a test dataset for our model.

@@

As can be seen from Figure 2, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap. This rotation is caused by the dimer realigning itself so that the gradient forces are minimised. Its interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely. In Vigilante's work with dimers \cite{5} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.

The data from Figure 2a was used to test the efficacy of our predictive model. By recording how much our estimate diverges from the true orientation we get a rough score of how well our model can predict the dimer's orientation.

@@

Large spikes in the divergence plot indicate that the model is deviating far from the desired prediction. From Figure 3 we can see our model has short bursts of being highly accurate in its prediction before suddenly shooting off. This can be in part be due to the fact that our $y(\hat{s}_k)$ values are not set but fluctuate due to the signal error. Reduction of the signal error results in closer and closer

>>>>

This has already been achieved with single microsphere using dynamic light scattering measurements of the trapping beam; where Bar-Ziv et al utilised dynamic light scattering to characterise the trap's stiffness and anisotropy based on the Brownian motion of a trapped silica microsphere \cite{1}. They note that their method assumes that the silica scatters isotropically which allows them to place a single detection fibre at $75^\circ$ from the propagation axis. In the case where the trapped object scatters anisotropically - such as for an asymmetrical dimer - it is necessary to utilise multiple detection fibres to build a better picture of the trapped object's orientation based on the produced light scattering. However, for inverted microscope optical traps there is a significant engineering challenge of placing enough detection fibres at the correct angles within a close enough distance to the trapping focus to collect a clear signal. This means that any fibre based detection will have to rely on maximising the information provided by the light scattering signals collected at a limited number of detection fibres. Excluding the engineering challenge involved with ensuring proper coupling between the probe laser and the detection fibres, the primary concern is how much useful information can be extracted from the trapped particle.

The experiemental set up that this is based on is a inverted microscope optical trap, a suspension of microparticles are loaded onto a plastic cover slide. Onto the slide four optical fibres are stuck down; one fibre walks a 633 nm HeNe laser onto the cover slip pointed directly at the trap focus, and the 3 remaining fibres are spaced equidistantly from the trap focus. The distance between each fibre and the trap focus is roughly $200\mu m$, with the exact angles between the particle and the trap focus being measured via the microcopes camera. The trapping laser is a 1064 nm NIR laser that is introduced via a 1.25 NA objective; to increase the trapping depth of the objective a lense of focal length 100 cm was added into the beam path to transform the profile entering the objective from parallel to converging, as advised by []'s paper.

@@

\section*{Results}
\subparagraph*{Asymmetric dimer dynamics}
The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap. The initial orientation was assumed as strictly vertical (in line with the beam propagation direction). The dimer's position orientation was recorded every $10 \mu s$ for using as a test dataset for our model.

@@

As can be seen from Figure 2, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap. Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis. It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely. Further simulations of differently sized dimers showed similar results, but only when $a_1 \geq 2a_2$. Dimers with more symmetrical size ratios immediately aligned into a fixed vertical position. 
In Vigilante's work with dimers \cite{5} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.

\subparagraph*{Predictive efficacy of model}
The data from Figure 2a was used to test the efficacy of our predictive model. By recording how much our estimate diverges from the true orientation we get a rough score of how well our model can predict the dimer's orientation.

@@

Spikes in the divergence plot indicate that the model is deviating  from the desired prediction. From Figure 3 we can see our model has short bursts of being highly accurate in its prediction before suddenly shooting off. This can be in part be due to the fact that our $y(\hat{s}_k)$ values are not set but fluctuate due to the signal error. Reduction of the signal error results in a higher accuracy in the models predictive efficacy to an extent; in truth the most significant factor in the prediction is our priori estimate of the dimer orientation $p(\hat{n})$. Since our choice of reference orientations have mirror copies in the physical space the conditional probabilities $p(y_k\parallel\hat{n})$ can be very close to each other. The addition of $p(\hat{n})$ means results in a significant improvement in the prediction by restraining our result to orientations close together in the physical space.

\subsection*{Further improvements}
While our model provides a satisfactory prediction of the dimer's motion; ideally this model can be exteneded to any particle characteristic that has a clear effect on the scattering from the trapped entity. For example using the mstm package we can determine how differently sized particles in similar orientations will produce different scattering patterns.

\section*{Conclusion}
We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. The
